---
title: "Assignment Tidydata& gganimate"
author: "Hank Chen"
date: "date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(tidyverse)
library(knitr)
library(kableExtra)
library(lubridate)
library(ggplot2)
library(gganimate) #this will runinto error on my computer...
```

# Your turn with tidyr

Let's work with a data set from one of your classmates.  

* Load the UN data into Rmarkdown file on canvas
* View this data 
* Which format is it? 
* What will be the ideal format for you to analyse this data in your project
* Change the format to that ideal format 
* Make a table like the one below using `kablestyling()`

```{r}
# call data set 
# change ito adequate format
UN<-read.csv("UN_ExampleTidy.csv")
UN1<-rename(UN, "Population"="Value", "Country"="Country.or.Area")
UN1<-select(UN1,Country,Population,Subgroup)
UN1<-head(spread(UN1,key="Subgroup", value="Population"),10)
UN1<-rename(UN1,"Female>15"="Female 15+ yr", "Total>15"="Total 15+ yr")
UN1<-subset(UN1,select=c(1,3,2))
UN1<-kable(UN1,"html")%>%
  kable_styling(full_width=FALSE, c("stripped","bordered", "hold_position"))%>%
  add_header_above(c(" "=1, "Population"=2))

UN1
```

# Your turn with ggnanimate

Find a time series dataset different from the ones used in class and make two cool animated graphs using gganimate. Have fun. 

<br>

```{r}
library(readxl)
library(gifski) 
Data <- read_excel("MyDataPart2.xls", 
    range = "A11:B752", col_types = c("date", "numeric"))

Data <- Data %>%
  separate(observation_date, sep="-", into = c("year", "month", "day"))

Data<-Data %>% group_by(year) %>% summarize(year_average=mean(CPILFESL))

ggplot(data=Data, mapping=aes(x=year, y=year_average)) +
  geom_point()+geom_smooth(method='lm')+scale_x_discrete(breaks=seq(1957, 2018, 10))+labs(title="Annual Consumer Price Index for All Urban Consumers")

# Code for animation but don't know if it's right because gganimate won't work on my computer

Data$year<-as.numeric(Data$year)

static<-ggplot(Data, aes(x=year, y=year_average)) + 
  geom_line(color="red")
static +
  transition_reveal(year) #reveal by date


# make y-axis to months and draw each months animation by years
Data2 <- read_excel("MyDataPart2.xls", 
    range = "A11:B752", col_types = c("date", "numeric"))
Data2 <- Data2 %>%
  separate(observation_date, sep="-", into = c("year", "month", "day"))
Data2<-Data2 %>% group_by(year)

Data2$year<-as.numeric(Data2$year)
Data2$month<-as.numeric(Data2$month)

monthly<-ggplot(data=Data2, mapping=aes(x=year, y=CPILFESL))+ geom_point(alpha=0.5)+facet_wrap(~month)+ylab("Monthly CPILFESL")+geom_line(color="blue")

monthly +
  transition_reveal(year) #reveal by date
```

