---
title: "Probability Functions and Number Generation"
author: "Hank Chen"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  prettydoc::html_pretty:
    theme: architect
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
#install.packages(tidyverse) remember this package should be already install. Packages are installed in this chunk so 
library(tidyverse)

```
# Probability Distribution Functions in R.  

Every distribution that R handles has four functions. There is a root name, for example, the root name for the normal distribution is norm. This root is prefixed by one of the letters

* p for "probability", the cumulative distribution function (c. d. f.)
* q for "quantile", the inverse c. d. f.
* d for "density", the density function (p. f. or p. d. f.)
* r for "random", a random variable having the specified distribution


Distributions | Functions 
--------------|----------
Binomial |	pbinom	qbinom	dbinom	rbinom
Uniform	| punif	qunif	dunif	runif
Normal |	pnorm	qnorm	dnorm	rnorm
Geometric	| pgeom	qgeom	dgeom	rgeom
Hypergeometric |	phyper	qhyper	dhyper	rhyper
Chi-Square |	pchisq	qchisq	dchisq	rchisq
Exponential |	pexp	qexp	dexp	rexp
F  | 	pf	qf	df	rf
Student  | t	pt	qt	dt	rt
Negative | Binomial	pnbinom	qnbinom	dnbinom	rnbinom
Poisson	| ppois	qpois	dpois	rpois


## Random number generation

In class you have learned about different probability distributions.
You can draw random numbers from existing distributions. It means that you can generate a set of random numbers that follow a specific distribution function.  Using R as the root and the abreviation for the specific distribution you can generate the random numbers, see examples below. 

## The Uniform Distribution
Were every value between a specified minimun and a maximun has an equal probability of ocurrence. the function in R that allows you to do that is `runif(n, min, max)`
```{r}
set.seed(123)
runif(1)

# Get a vector of 4 numbers
runif(4)

# Get a vector of 3 numbers from 0 to 100
runif(3, min=0, max=100)

# Get 5 integers from 0 to 100
# Use max=101 because it will never actually equal 101
data1<-floor(runif(5, min=0, max=101))
# look at the object data in the enviroment
```

The sample function also generates random numbers.
```{r}
# This will do the same thing as the last code above, WITH replacement. 
# sample(1:100, 10, replace=TRUE)

# To generate integers WITHOUT replacement:
# sample(1:100, 10, replace=FALSE)

# You can also create a sample out a data set. 
# sample(data1, 10, replace=FALSE)
```

## The Normal distribution 
To generate numbers from a normal distribution, use rnorm(). 

```{r}
# By default the mean is 0 and the standard deviation is 1.
set.seed(45)
rnorm(4)

# Use a different mean and standard deviation
rnorm(4, mean=50, sd=10)

```

# More graphs with ggplot 

Two type of graphs allow you to show the distribution of a variable.

##  Histogram and Bar graph 

A histogram is bar graph that shows the frequency of occurence of each value. 

```{r}
# To check that the distribution looks right, make a histogram of the numbers
set.seed(45)
x <- rnorm(400, mean=50, sd=10)
# Using the basic R graphs hist() create a histogram
hist(x)
# Using qplot 
qplot(x, geom="histogram")

y <- rnorm(400, mean=5, sd=0.5)

# To use ggplot you need your data in a data frame structure, that is the only way to feed data into ggplot.
dat<-data.frame(x,y)

```


```{r}
# A bar graph in ggplot 
ggplot(dat, aes(x))+
  geom_bar(stat='bin', fill="green", alpha=0.5)+
labs(title="Histogram of X")+ theme(plot.title = element_text(hjust = 0.5))

# A histogram in ggplot 
ggplot(dat, aes(x))+
  geom_histogram(fill="green", alpha=0.5)+
labs(title="Histogram of X")+ theme(plot.title = element_text(hjust = 0.5))

#See they look the same
```
# Making a better histogram
```{r}

# Once you create your graph you see a warning about the bins change that using breaks=seq()
# Here you also change the y axis to be the density instead of the frequency. 
ggplot(dat, aes(x))+
geom_histogram(aes(y =..density..), fill="green", breaks=seq(20, 80, by = 2), col="red", alpha = .2)+ geom_density(col=2) + 
labs(title="Histogram of X")+ theme(plot.title = element_text(hjust = 0.5))

#Now you want to change the attribute fill= so that it incorporates de information about the number (count) of observation in that range.
#Here you are also removing the atribute color from the geom_hist adding the geom_density() with the color attribute being equal to 3 
ggplot(dat, aes(x))+
geom_histogram(aes(y =..density.., fill=..count..), breaks=seq(20, 80, by = 2), alpha = .8)+ geom_density(col=3) + labs(title="Histogram of X")+ theme(plot.title = element_text(hjust = 0.5))
```

# Boxplot 

A boxplot is also a graph that provides a visualization about the distribution of your variable of vector 

Below a boxplot for the avg_SAT in our college data. 

```{r}
college <- read.csv("http://672258.youcanlearnit.net/college.csv")  

# Box plot
ggplot(data=college, aes(x="",  y=sat_avg))+
  geom_boxplot(fill="green", alpha=0.5)+
labs(title="SAT Boxplot") +
   theme(plot.title = element_text(hjust = 0.5))


# Boxplot that incorporates  separates the sat in public and private showing the distribution and also assigns the control variable to the atribute fill so each on has a different color. 
ggplot(data=college, aes(x=control,  y=sat_avg, fill=control))+
  geom_boxplot( alpha=0.5)+
labs(title="SAT Boxplot") +
   theme(plot.title = element_text(hjust = 0.5))

ggplot(data=college, aes(x=control,  y=tuition, color=control))+
  geom_boxplot()+
labs(title="SAT Boxplot") +
   theme(plot.title = element_text(hjust = 0.5))


```


# Your turn Assingment 2 

Play around with this functions, start by looking at the documentation in R. For example ?punif  or ?qnorm ? sample 

* Create a new Rmardown document 

* Generate a vector y of 100 observations from the normal distribution with mean=3 and standart deviation sd=7 

* Generate a vector z of 100 observations from the normal distribution with mean=7 and sd=3

* Generate a vector u of 100 observations from the uniform distribution 
* Generate a vector v of 100 observations of integers between 0 to 500 without replacement 
* Make a histogram of y and v 
* Make a box plot of z 
* Using the data from college make a boxplot of the average faculty salary separate by public and private universities

```{r}
set.seed(245)
```



