---
title: "Lab6 Descriptives with dplyr and Cleaning Data Editing"
author: "Dr. Paloma L. Moyano"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    theme: readable
    toc: yes
  word_document:
     toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
library(tidyverse)
```

# 1 Probabilities of an event

The function `choose(n, k)` in R allows you to choose k elements out of a set of n. 
For example: 
*Choose 7 apples out a bag of 10 apples 5 red 5 green 
*Event G: Choose exactly 2 green apples 

```{r}
n_apples<-choose(10,7)
n_g<-choose(5,3)*choose(5,2)
n_apples
n_g
```

Let's practice 
* Define an experiment: Choose any five representatives out of 20 where 8 are males and 12 are females. Name this object n_s that contains those 5 representatives. 
* Define event A: Choose exactly three female representatives create an object n_a that represens event A 
* Calculate the probability of A happening, name it prob_a

To create the chunk of code Ctr+Alt+i and start writing your code

```{r}
(n_s<-choose(20,5))
(n_a<-choose(12,3)*choose(8,2))
(prob_a <- n_a/n_s)
```


# 2 Compute mean, median, variance and correlation 

Let's say you have two variables `j` and `k` the table below tells you the functions in R to calculate these descriptive statistics

Statistic | Function in R | Alternatives
------------- | ------------- | --------
Mean of `j`  | mean(j) | sum(j)/lenght(j)
Median of `j` | median(j) |
Variance of `j` | var(j) | sum((j - mean(j))^2)/length(j) | var(j)*((length(j)-1)/lenght(j))
Standard deviation of `j` | sd(j) | sqrt(var(j))
Correlation of `j` and `k` | cor(j, k) |
Covariance of `j` and `k` | cov( j, k) |


Remember you can use the package the %>% option in the package dplyr to manipulate your data and obtain these using summarise and summarise_all 

```{r, eval=FALSE}
df # lest say you have a dataframe called df
df %>% summarise_all(funs(mean, var)) # this takes all the variables in your dataframe and applies the functions mean and variance to them 
mean(df$variable) # this will give you the mean of the specific variable in your dataframe 
```

For the following examples we will use data from DASL about the fat content of different burger brands, this data is in text and we can load it into R using the code below.

```{r}
burgers <- read.table("https://dasl.datadescription.com/download/data/3088.txt", header = TRUE)
```

* Compute the mean of all variables. 

* Find the correlation between Fat and Sodium. Name the object cor_fs

* Compute the standard deviation of Calories. Name the object sdCal

* Calculate the variance of Sodium and Calories and the covariance of those two

Create your chunk of code here
```{r, }
burgers %>% summarise_all(mean)
burgers %>% summarise(cor_fs=cor(Fat, Sodium), sdCal=sd(Calories), var(Sodium), var(Calories), cov(Sodium, Calories))
```


Let's look at an exmaple with F2019 student data

```{r}

load("~/Desktop/EMORY2019Fall/econ220/Econ220DataF2019_ano.Rdata")
# Let's give it another name so it is easier to work with it.
Prodata<-Econ220DataF2019_ano

#cleaning the q182 variable about travelling
class(Prodata$q182)
data.frame(table(Prodata$q182))
Prodata$q182[Prodata$q182=="1 time"]<-1
Prodata$q182[Prodata$q182=="2-3"]<-2.5
Prodata$q182[Prodata$q182=="2-3 times"]<-2.5
Prodata$q182[Prodata$q182=="3 times"]<-3
Prodata$q182[Prodata$q182=="4 times"]<-4
Prodata$q182[Prodata$q182=="5+"]<-5
Prodata$q182[Prodata$q182=="once per semester"]<-1
Prodata$q182[Prodata$q182=="Three"]<-3
Prodata$q182[Prodata$q182=="twice"]<-2
Prodata$q182 <- as.numeric(Prodata$q182,na.rm = TRUE)
summary(Prodata$q182)

Prodata
Prodata<-Prodata %>% filter(!is.na(year))
table(Prodata$year)
Prodata$year<-ordered(Prodata$year, levels=c("Freshmen","Sophomore","Junior","Senior"))
Prodata  %>% group_by(year)%>% summarize(mean(q182, na.rm=TRUE))
levels(Prodata$year)
class(Prodata$q182)
mean(Prodata$q182)
Prodata %>% filter(!is.na(year))  %>% group_by(year)%>% summarize(averagetravel=round(mean(q182, na.rm=TRUE),3))

data.frame(table(Prodata$q15))

# Editing string in 
#sub(old, new, string)
#gsub(old, new, string)

Prodata$q15<-gsub("^<|\\hour|\\hours|\\h|\\minutes", "", Prodata$q15) 
data.frame(table(Prodata$q15))
Prodata$q15<-as.numeric(Prodata$q15)
data.frame(table(Prodata$q15))
Prodata$q15[Prodata$q15==40]<-40/60
data.frame(table(Prodata$q15))

#A function to replace strings. 
#str_replace_all(string, pattern, replacement)
# simple example 
fruits <- c("one apple", "two pears", "three bananas")
fruits %>%
  str_replace_all(c("one" = "1", "two" = "2", "three" = "3"))

# Let's do it on variable for major
Prodata$major<-str_replace_all(Prodata$major, "Economics", "THE BEST")
#View(Prodata$major)
```

