---
title: "tidyverse and dplyr package"
author: "Dr. Paloma L. Moyano"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  html_document: 
    theme: yeti
    toc: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

## Installing packages
R has many "packages", essentially add-ons to the basic R program that allow you to implement new techniques or access new methods.  One absolutely essential add-on is called `tidyverse`, which provides a set of methods for manipulating data and visualizing it. The package babynames has cool data on the babynames for the US that we will use in this class too. 

A package is like an app on your phone, you install it but everytime you need it you have to open it, in the case of packages in R you install them and in order to call them everytime you use the function `library(packagename)`.
```{r, echo = T, message = F, warning = F, results='hold'}
install.packages("tidyverse")
install.packages("babynames")
library(tidyverse)
library(babynames)
```
Today we will simply load some pre-built data from the `babynames` package, and use the `str()` and `summary()` functions understand the content of the data 
```{r, eco = T}
data(babynames, package="babynames") # loads the data 
str(babynames) # provides the structure of the data.
summary(babynames) # provides you a good summary of each variable in the data set 
```
The data we loaded is from the baby name data for each year from 1880 to 2015, the number of children of each sex given each name. All names used 5 or more times are included. As you can see from the summary, the data set contains five variables:year, sex, name, n the number of times the name was used, prop de proportion of times the name was used.  

# Data Manipulation with `dplyr`

## `dplyr` functions

function |  Purpose
-----------|--------
 filter()   | Select rows based on criteria
 arrange()  | order rows
 select()   | choose columns
 distinct() | unique values of column(s)
 mutate() and transmute() | Create new columns
 summarise()  | Find a summary statistic
 sample_n() | Take a random sample
 group_by() | Creates groups based on the values or categories of a variable


### Understanding what `%>%` does

You most likely have noticed each of these `dplyr` examples involve `%>%`.  This expression roughly translates to "and then".  So for example 
```{r, echo = T, eval = F}
babynames %>% filter(sex == "M" & year==1997) %>% head()
```
Roughly translates to: Take the babynames dataset **and then** keep only those observations that are male **and then** show the first 6 observations of that data set. If you execute this code, the output, shown above, will be the first 6 observations of a dataset that includes only men born in 1997.
 
### Using `filter()` to select observations

Let's practice what you learned. Create a new data set with just boys, name it just_boys. You could also create a dataset of just girls born between 1997 and 2002, name this dataset girls21

```{r, echo = T}
just_boys<- babynames %>% filter(sex=="M")
girls21<-babynames %>% filter(sex=="F", year>=1997 & year<2002) %>% head()
just_boys<- babynames %>% filter(sex=="M")
girls21big<-babynames %>% filter(sex=="F", year>=1997 & year<2002)
girls21
girls21big
#head(girls21,10)
```

### Using `arrange()` to order observations
You can order the observations by a column, lets check 
```{r, echo = T}
youngest_first <- babynames %>% arrange(desc(year))
head(youngest_first)
tail(youngest_first)
```

### Your turn: combine filter with arrange. 
Find the 6 most popular names for your gender in the year you where born. 

```{r, echo = T}
babynames %>% filter(sex=="F", year==1999)%>%arrange(desc(n))%>% head()

```

### Using `mutate()` to create new columns (variables) in a dataset

Suppose hypothetically you wanted to create a new variable called "popular" when the names where chosen more that 10 thousand times. To create a new dataset that included this variable, you would have to use `mutate()`

```{r, echo = T}
new_babynames <- babynames %>% mutate("popular" = n>10000)
head(new_babynames)
```

### Using `select()` to select columns

The `select()` function is designed to allow one to select specific columns of the dataset.
Below different ways of making the same selection in R. 

```{r, echo=T}
no_prop <-babynames %>% select(year, sex, name, n)
head(no_prop)
no_prop1 <- babynames %>% select(-prop)
head(no_prop1)
no_prop2 <- babynames %>% select(starts_with("pr"))
no_prop3 <- babynames %>% select(year : n )
head(no_prop3)
```

### Using `summarise()` to find summary statistics

Suppose we wanted to find the average popularity of names, the number of observations in the data set, and maximun popularity in the same year. 

```{r, echo = T}
babynames %>% summarise("Avg_popularity" = mean(n), 
                      "N" = length(n), 
                      "Most_popular" =  max(n)) 

filter(babynames, n==max(n))
```

### Finding statistics by group using `group_by()`

The `group_by()` functions divides the data into groups for all future commands.  This can be extremely useful in combination with the `summarise()` function.  For example, suppose we wanted to find the average popularity of a name by sex in all this time:

```{r, echo = T}
babynames %>% group_by(sex) %>% summarise("Avg_pop"= mean(n))
```
What if you wanted to divide the data into groups by both year and sex we can just use the variables separated by a comma in the `group_by()` function. 


##In class exercise

Find the names that were most popular female and male in the year you were born. The verb top_n() gives you the top n observations based on the variable you use for ordering, in the console do ?top_n to see the arguments of this funtion and use it in the exercise.   

```{r, echo = T}

```








