---
title: "Assignment 2"
author: "Hank Chen"
date: "10/27/2019"
output: 
  prettydoc::html_pretty:
    theme: architect
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
#install.packages(tidyverse) remember this package should be already install. Packages are installed in this chunk so 
library(tidyverse)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
# Your turn Assingment 2 

Play around with this functions, start by looking at the documentation in R. For example ?punif  or ?qnorm ? sample 

* Create a new Rmardown document 

* Generate a vector y of 100 observations from the normal distribution with mean=3 and standart deviation sd=7 

* Generate a vector z of 100 observations from the normal distribution with mean=7 and sd=3

* Generate a vector u of 100 observations from the uniform distribution 
* Generate a vector v of 100 observations of integers between 0 to 500 without replacement 
* Make a histogram of y and v 
* Make a box plot of z 
* Using the data from college make a boxplot of the average faculty salary separate by public and private universities

```{r}
set.seed(245)

y <- rnorm(100, mean=3, sd=7)
z <- rnorm(100, mean=7, sd=3)
u <- runif(100)
v <- sample(1:500, 100, replace=FALSE)

# Store y, v and z into a data frame before graphing
dat<-data.frame(y,v,z)

# Make a histogram of y
ggplot(dat, aes(y))+
geom_histogram(aes(y =..density.., fill=..count..), breaks=seq(-10, 20, by = 2), alpha = .8)+ geom_density(col=3) + labs(title="Histogram of Y")+ theme(plot.title = element_text(hjust = 0.5))

# Make a histogram of v
ggplot(dat, aes(v))+
geom_histogram(aes(v, y =..density.., fill=..count..), breaks=seq(10, 500, by = 20),col="red", alpha = .8)+ geom_density(col=2) + labs(title="Histogram of V")+ theme(plot.title = element_text(hjust = 0.5))

# Make a box plot of z 
dat2<-data.frame(z)
ggplot(data=dat2, aes(x="",  y=z))+
  geom_boxplot(fill="green", alpha=0.5)+
labs(title="z Boxplot") +
   theme(plot.title = element_text(hjust = 0.5))


college <- read.csv("http://672258.youcanlearnit.net/college.csv")  

# Using the data from college to make a boxplot of the average faculty salary separate by public and private universities
ggplot(data=college, aes(x=control,  y=faculty_salary_avg, fill=control))+
  geom_boxplot( alpha=0.5)+
labs(title="SAT Boxplot") +
   theme(plot.title = element_text(hjust = 0.5))

# Using the data from college to make a boxplot of the median debt separate by public and private universities
ggplot(data=college, aes(x=control,  y=median_debt, color=control))+
  geom_boxplot()+
labs(title="SAT Boxplot") +
   theme(plot.title = element_text(hjust = 0.5))

```