---
title: "Cleaning Project data"
author: "Paloma Lopezdemesa"
date: "10/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Cleaning project data 

```{r}
# Load data into R, the data is already in R format
load("~/Desktop/EMORY2019Fall/econ220/Econ220DataF2019_ano.Rdata")
# Let's give it another name so it is easier to work with it.
Prodata<-Econ220DataF2019_ano
```

# Data editing and cleaning 

Unfortunately some of you gave answers that afected the class of the variables or that had more detail than necesary and that needs edits. 
Please do reveal who you are in class unless you want studnets to see your personal data can see your personal data. 

```{r, eval=FALSE}
# If you want to save the data in csv because you want to see it in #excel this is the code you use to save into csv. However, I 
#discourage you to use excel to edit your data, because it is not easy #to keep record of the edits. 
write.csv(Prodata, "~/Dropbox/Prodata.csv", row.names=FALSE)
```

Let's look for example at Question 38 in the code book. 
*What is the hourly wage of your current job in college?*
*Give a numerical answer leave blank if you don t have a job*
Supose you are interested in investigating this question, first you need to make sure the answers make sense.
```{r}
#View(Prodata$q38) allow you to see how this variable looks
# Ideally this variable is numeric but because of the answers 
#it is chr variable. use class to see this
class(Prodata$q38)
# You can also look at the variable specific cases that you need 
#to fix and proceed to edit the variable as needed. 
data.frame(table(Prodata$q38))
# Now the edits
Prodata$q38[Prodata$q38=="$8.00"]<-"8"
Prodata$q38[Prodata$q38=="0(can't leave this blank, it is a point for the survey)"]<-NA
Prodata$q38[Prodata$q38=="8.50 workstudy, 17/hr online chess tutoring"]<-8.5
# last edit: making this variable numeric
Prodata$q38<-as.numeric(Prodata$q38)
# Check that the change was made
class(Prodata$q38)
# Fequency table again to make sure everything looks good now
summary(Prodata$q38)
```

#In class exercise 
Fix question 17 in the Data
*What was your high school GPA ?*

```{r}


```

